---
title: "Alternative management"
author: "Marlese"
date: "2/9/2021"
output: html_document
---


# opsies

```{r}

# chi square
# kruskal wallis
# fisher exact test
# ordered logistic regression
# multi discriminant analysis
# decision trees


```


# librarires

```{r}

library(openxlsx)
library(readr)

library(tidytext)
library(tidyr)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(rvest)
library(xml2)
library(dplyr)
library(plyr)

library(MASS)  #ordered logistic regression

# logistic
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)


library(nnet)

library(wesanderson)

```

#import

```{r}

manage_raw = read.xlsx("C:\\Users\\2010088819\\Documents\\R Projects\\Scle_management\\manage.xlsx", "temp")


```

# temperature
## modify

```{r}



manage_raw$size_mm <- as.factor(manage_raw$size_mm)

manage_raw$weight_g <- as.factor(manage_raw$weight_g)

manage_raw$rep <- as.factor(manage_raw$rep)

manage_raw$depth_cm <- as.factor(manage_raw$depth_cm)

manage_raw$time_min <- as.factor(manage_raw$time_min)

manage_raw$temp <- as.factor(manage_raw$temp)

manage_raw$myc <- as.character(manage_raw$myc)

manage_raw$scle <- as.character(manage_raw$scle)




manage_raw1 <- manage_raw %>%
    filter(temp %in% c("125", "155", "185", "200")) %>% 
    mutate(myc1 = case_when(myc == "0" ~ "absent",
                            myc == "1" ~ "contamination",
                            myc == "2" ~ "present_partial",
                            myc == "3" ~ "present_full",
                            myc != "*" ~ myc),
           scle1 = case_when(scle == "0" ~ "absent",
                             scle == "1" ~ "contamination",
                             scle == "2" ~ "present_partial",
                             scle == "3" ~ "present_full",
                             scle != "*" ~ scle)) 
  




#Ordering the  variable, maar is nie werklik in orders nie
# manage_raw$myc1 = factor(manage_raw$myc1, levels = c("absent", "contamination", "present_partial", "present_full"), ordered = TRUE)
# manage_raw$scle1 = factor(manage_raw$scle1, levels = c("absent", "contamination", "present_partial", "present_full"), ordered = TRUE)
# manage_raw$temp = factor(manage_raw$temp, levels = c("125", "155", "185", "200"), ordered = TRUE)
# manage_raw$size_mm = factor(manage_raw$size_mm, levels = c("15", "25", "35", "45"), ordered = TRUE)
# manage_raw$depth_cm = factor(manage_raw$depth_cm, levels = c("0", "5"), ordered = TRUE)
# manage_raw$time_min = factor(manage_raw$time_min, levels = c("5", "10", "15"), ordered = TRUE)






```

## statistical analysis tets

### chi square

<!-- ```{r} -->

<!-- chisq.test(manage_raw$myc1, manage_raw$time_min, correct=FALSE) -->

<!-- ``` -->

### kruskal wallis

<!-- ```{r} -->

<!-- #Ordering the  variable, maar is nie werklik in orders nie -->
<!-- manage_raw$myc1 = factor(manage_raw$myc1, levels = c("absent", "contamination", "present_partial", "present_full"), ordered = TRUE) -->
<!-- manage_raw$scle1 = factor(manage_raw$scle1, levels = c("absent", "contamination", "present_partial", "present_full"), ordered = TRUE) -->


<!-- kruskal.test() -->

<!-- ``` -->


### fisher

<!-- ```{r} -->


<!-- fisher.test(manage_raw1) -->


<!-- ``` -->




## discriminant analysis

Compared to logistic regression, the discriminant analysis is more suitable for predicting the category of an observation in the situation where the outcome variable contains more than two classes. Additionally, it’s more stable than the logistic regression for multi-class classification problems(http://www.sthda.com/english/articles/36-classification-methods-essentials/146-discriminant-analysis-essentials-in-r/)

### ordered logistic regression 

<!-- ## description -->
<!-- ### table1 -->
<!-- one at a time, table elke variable -->

<!-- ```{r} -->

<!-- #Ordering the  variable, maar is nie werklik in orders nie -->

<!-- manage_raw1$myc1 = factor(manage_raw1$myc1) -->
<!-- manage_raw1$scle1 = factor(manage_raw1$scle1) -->

<!-- head(manage_raw1) -->

<!-- summary(manage_raw1) -->

<!-- lapply(manage_raw1[, c("myc1", "temp", "time_min", "size_mm", "depth_cm")], table) -->


<!-- ``` -->

<!-- ###table2 -->

<!-- five way cross tabs (xtabs) and flatten the table -->

<!-- ```{r} -->

<!-- ftable(xtabs(~ size_mm + myc1 + temp + depth_cm + time_min, data = manage_raw1)) -->

<!-- ``` -->


### logistic regression

<!-- nie manier om te gaan nie. Te min data punte -->
<!-- ```{r} -->



<!-- m <- polr(myc1 ~ time_min + temp + depth_cm + size_mm, data = manage_raw1, Hess=TRUE) -->

<!-- s <- polr(scle1 ~ time_min + temp + depth_cm + size_mm, data = manage_raw1, Hess=TRUE) -->

<!-- summary(m) -->

<!-- summary(s) -->

<!-- ``` -->


<!-- ### p value -->

<!-- ```{r} -->

<!-- # store coeff table -->
<!-- (ctable <- coef(summary(m))) -->

<!-- # calculate and store p values -->
<!-- p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2 -->

<!-- # combined table -->
<!-- (ctable <- cbind(ctable, "p value" = p)) -->


<!-- ``` -->

<!-- ### confidence intervals -->

<!-- Onthou die is in log formaat -->
<!-- #### logs -->
<!-- ```{r} -->

<!-- #profiled CI (nie simmetries) -->

<!-- ci <- confint(m) -->

<!-- #CI assume normality -->
<!-- # as 97.5 nie verby 0 gaan nie is dit significant -->
<!-- confint.default(m) -->

<!-- ``` -->

<!-- #### exsp -->
<!-- Another way to interpret logistic regression models is to convert the coefficients into odds ratios. -->
<!-- ```{r} -->

<!-- ## odds ratios -->
<!-- exp(coef(m)) -->

<!-- ## OR and CI -->
<!-- ### genoem: proportional odds ratios -->
<!-- exp(cbind(OR = coef(m), ci)) -->

<!-- ``` -->

### multinomial logistic regression

<!-- This is the regression you use when you have an unordered categorical variable outcome or an ordered categorical variable outcome that failed the Brant test (https://rpubs.com/rslbliss/r_logistic_ws) -->

<!-- https://stats.idre.ucla.edu/r/dae/multinomial-logistic-regression/ -->

<!-- hoekom die dalk nie die regte model is nie is agv die klein sample size -->

<!-- ```{r} -->

<!-- manage_raw1$myc1 = factor(manage_raw1$myc1) -->
<!-- manage_raw1$scle1 = factor(manage_raw1$scle1)  -->


<!-- #kyk na data -->
<!-- with(manage_raw1, table(time_min, temp, size_mm, depth_cm, myc1)) -->

<!-- # fit model -->
<!-- model_multi <- multinom(myc1 ~ time_min + temp + size_mm + depth_cm, data = manage_raw1) -->


<!-- summary(model_multi) -->

<!-- #p value stappe -->
<!-- z <- summary(model_multi)$coefficients/summary(model_multi)$standard.errors -->
<!-- z -->


<!-- p <- pnorm(abs(z), lower.tail=FALSE)*2 -->
<!-- p -->

<!-- ## extract the coefficients from the model and exponentiate -->
<!-- exp(coef(model_multi)) -->

<!-- #calculate predicted propabilitites -->
<!-- head(pp <- fitted(model_multi)) -->

<!-- #model performance - acuuracy -->

<!-- most_likely <- apply(model_multi$fitted.values, 1, which.max) ## Applies the which.max function to each row of the matrix of predicted probabilities -->

<!-- pred_cats <- levels(manage_raw$myc1)[most_likely]  ## Map the maximums (either 1, 2, or 3) to category labels -->

<!-- sum(pred_cats == manage_raw$myc1)/nrow(manage_raw1) ## Calculate the fraction of correct predictions (ie: accuracy) -->



<!-- ``` -->

<!-- ### classification trees -->

<!-- https://remiller1450.github.io/s230f19/caret3.html -->

<!-- kort ook 'n min van 10 punte per variable -->

<!-- ```{r} -->

<!-- library(rpart) -->
<!-- library(nnet) -->
<!-- library(pROC) -->
<!-- library(rpart) -->
<!-- library(rpart.plot) -->

<!-- #create -->
<!-- model_tree <- rpart(myc1 ~ temp + time_min , manage_raw1) -->
<!-- prp(model_tree, extra = 104) -->


<!-- ggplot(manage_raw1, aes(x = temp, y = time_min, color = myc1, pch = myc1)) +  -->
<!--   geom_point() +  -->
<!--   geom_vline(xintercept = 155, lty = 2, lwd = 1.5)  -->

<!-- p1 <- ggplot(manage_raw1, aes(x = temp, y = time_min, color = myc1, pch = myc1)) + geom_point() + geom_vline(xintercept = 125, lty = 2, lwd = 1.5) + geom_segment(aes(x = min(temp-1), y = 10, xend = 125, yend = 10), lwd = 1.5, lty  =2, color = "black") + labs(title = "Splitting the left node") + theme(legend.position = "none") -->

<!-- p2 <- ggplot(manage_raw1, aes(x = temp, y = time_min, color = myc1, pch = myc1)) + geom_point() + geom_vline(xintercept = 155, lty = 2, lwd = 1.5) + geom_segment(aes(x = min(temp-1), y = 10, xend = 155, yend = 10), lwd = 1.5, lty  =2, color = "black") + geom_segment(aes(x = 185, y = 1.98, xend = 185, yend = 4.45), lwd = 1.5, lty  =2, color = "black")+ labs(title = "Splitting the left and right nodes") + theme(legend.position = "none") -->

<!-- gridExtra::grid.arrange(p1, p2, nrow = 1) -->

<!-- ``` -->



## histogram

```{r}
freq_df <- manage_raw %>%
  filter(temp %in% c("125", "155", "185", "200")) %>% 
  mutate(binary = case_when(myc == 0 ~ 0,
                            myc == 1 ~ 0,
                            myc == 2 ~ 1,
                            myc == 3 ~ 1)) %>% 
  group_by(size_mm, weight_g, depth_cm, time_min, temp) %>% 
  dplyr::summarise(freq = sum(binary)/4*100)

new_label_time <- c("5" = "5 min", "10" = "10 min", "15" = "15 min")
new_label_depth <- c("0" = "0 cm", "5" = "5 cm")
new_label_weight <- c("0.054" = "0.054 g", "0.222" = "0.222 g", "0.67" = "0.670 g", "1.31" = "1.310 g")


freq_df %>% 
  ggplot()+
  geom_boxplot(aes(temp, freq), fill = "grey")+
  facet_grid(~weight_g~time_min, labeller = labeller(weight_g = new_label_weight, time_min = new_label_time))+
    theme_bw()+
    labs(x = expression(paste("Temperature (",degree,"C)")), y = "Germination (%)")+
  theme(axis.title.x=element_text(size=18,face="bold"),
       axis.title.y=element_text(size=18,face="bold"),
       axis.text.y = element_text(size = 14, color = "black", face = "bold"),
       axis.text.x = element_text(size = 14, color = "black", face = "bold"),
       legend.text = element_text(size = 16),
       legend.title = element_text(size = 18),
       strip.background = element_rect(fill="white", size=1, color="black"),
       strip.text.x = element_text(size=14, face="bold"),
       strip.text.y = element_text(size=14, face="bold"))+
         ggsave("plots/variables_weight.png", units = "cm", width = 60, height = 30, dpi = 300)

freq_df %>% 
  ggplot()+
  geom_boxplot(aes(temp, freq), fill = "grey")+
  facet_grid(~depth_cm~time_min, labeller = labeller(depth_cm = new_label_depth, time_min = new_label_time))+
    theme_bw()+
    labs(x = expression(paste("Temperature (",degree,"C)")), y = "Germination (%)")+
  theme(axis.title.x=element_text(size=18,face="bold"),
       axis.title.y=element_text(size=18,face="bold"),
       axis.text.y = element_text(size = 14, color = "black", face = "bold"),
       axis.text.x = element_text(size = 14, color = "black", face = "bold"),
       legend.text = element_text(size = 16),
       legend.title = element_text(size = 18),
       strip.background = element_rect(fill="white", size=1, color="black"),
       strip.text.x = element_text(size=14, face="bold"),
       strip.text.y = element_text(size=14, face="bold"))+
         ggsave("plots/variables_depth.png", units = "cm", width = 60, height = 30, dpi = 300)

  
  
library(ggpubr)

ggscatter(manage_raw, x = "size_mm", y = "weight_g",
          add = "reg.line",
          conf.int = TRUE,
          add.oarams= list(color = "blue",
                           fill = "gray"))+
  stat_cor(method = "pearson", label.x = 15, label.y = 1)
```






## anova

```{r}




aov.freq <- aov(freq ~ weight_g*temp + weight_g*time_min + weight_g*depth_cm + temp*time_min + temp*depth_cm + time_min*depth_cm + temp*time_min*depth_cm + temp*depth_cm*weight_g + time_min*depth_cm*weight_g,  freq_df)

summary(aov.freq)
anova(aov.freq)

write.xlsx(aov.freq, "anova_manage.xlsx")

#residual standard error

sqrt(deviance(aov.freq)/df.residual(aov.freq))

```

## LSD
http://agroninfotech.blogspot.com/2020/08/barplot-for-main-and-interaction.html

```{r}
library(agricolae)

a <- LSD.test(aov.freq, "temp", DFerror = aov.freq$df.residual, MSerror = deviance(aov.freq)/aov.freq$df.residual, alpha = 0.05, group = TRUE)
b <- LSD.test(aov.freq, "time_min", DFerror = aov.freq$df.residual, MSerror = deviance(aov.freq)/aov.freq$df.residual, alpha = 0.05, group = TRUE)
c <- LSD.test(aov.freq, "depth_cm", DFerror = aov.freq$df.residual, MSerror = deviance(aov.freq)/aov.freq$df.residual, alpha = 0.05, group = TRUE)
d <- LSD.test(aov.freq, "weight_g", DFerror = aov.freq$df.residual, MSerror = deviance(aov.freq)/aov.freq$df.residual, alpha = 0.05, group = TRUE)

ab = LSD.test(aov.freq,
                  c("temp", "time_min"),
                  DFerror = aov.freq$df.residual,
                  MSerror = deviance(aov.freq)/aov.freq$df.residual,
                  alpha = 0.05,
                  group = TRUE,
                  console = TRUE)

ac = LSD.test(aov.freq,
                  c("temp", "depth_cm"),
                  DFerror = aov.freq$df.residual,
                  MSerror = deviance(aov.freq)/aov.freq$df.residual,
                  alpha = 0.05,
                  group = TRUE,
                  console = TRUE)

bc = LSD.test(aov.freq,
                  c("time_min", "depth_cm"),
                  DFerror = aov.freq$df.residual,
                  MSerror = deviance(aov.freq)/aov.freq$df.residual,
                  alpha = 0.05,
                  group = TRUE,
                  console = TRUE)

ad = LSD.test(aov.freq,
                  c("temp", "weight_g"),
                  DFerror = aov.freq$df.residual,
                  MSerror = deviance(aov.freq)/aov.freq$df.residual,
                  alpha = 0.05,
                  group = TRUE,
                  console = TRUE)

abc = LSD.test(aov.freq,
                  c("temp", "time_min", "depth_cm"),
                  DFerror = aov.freq$df.residual,
                  MSerror = deviance(aov.freq)/aov.freq$df.residual,
                  alpha = 0.05,
                  group = TRUE,
                  console = TRUE)


a
b
c
d

ab
ac
bc
ad

abc
```
##### original grouping

```{r}
## Original order of LSD$group 

ascend_A = a$groups %>%
          group_by(rownames(a$groups)) %>%
          arrange(rownames(a$groups))

print(ascend_A)

ascend_B = b$groups %>%
           group_by(rownames(b$groups)) %>%
           arrange(rownames(b$groups))

print(ascend_B)


ascend_C = c$groups %>%
          group_by(rownames(c$groups)) %>%
          arrange(rownames(c$groups))

print(ascend_C)

ascend_D = d$groups %>%
          group_by(rownames(d$groups)) %>%
          arrange(rownames(d$groups))

print(ascend_D)

ascend_AB = ab$groups %>%
          group_by(rownames(ab$groups)) %>% 
          arrange(freq)

print(ascend_AB)

ascend_AC = ac$groups %>%
          group_by(rownames(ac$groups)) %>%
          arrange(freq)

print(ascend_AC)

ascend_BC = bc$groups %>%
          group_by(rownames(bc$groups)) %>%
          arrange(freq)

print(ascend_BC)

ascend_AD = ad$groups %>%
          group_by(rownames(ad$groups)) %>%
          arrange(freq)

print(ascend_AD)

ascend_ABC = abc$groups %>%
          group_by(rownames(abc$groups)) %>%
          arrange(freq)

print(ascend_ABC)


```


##### Mean error and DF

```{r}
meanSE_a = freq_df %>%
          dplyr::group_by(temp) %>% 
          dplyr::summarise(avg_a = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))

print(meanSE_a)

meanSE_b = freq_df %>%
          dplyr::group_by(time_min) %>%
          dplyr::summarise(avg_b = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))
print(meanSE_b)

meanSE_c = freq_df %>%
          dplyr::group_by(depth_cm) %>%
          dplyr::summarise(avg_c = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))
print(meanSE_c)

meanSE_d = freq_df %>%
          dplyr::group_by(weight_g) %>%
          dplyr::summarise(avg_d = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))
print(meanSE_d)

MeanSE_AB = freq_df %>%
          dplyr::group_by(temp, time_min) %>%
          dplyr::summarise(avg_AB = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))

MeanSE_AB = MeanSE_AB %>% 
  arrange(avg_AB)


print(MeanSE_AB)

MeanSE_AC = freq_df %>%
          dplyr::group_by(temp, depth_cm) %>%
          dplyr::summarise(avg_AC = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))

MeanSE_AC = MeanSE_AC %>% 
  arrange(avg_AC)

print(MeanSE_AC)

MeanSE_BC = freq_df %>%
          dplyr::group_by(time_min, depth_cm) %>%
          dplyr::summarise(avg_BC = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))

MeanSE_BC = MeanSE_BC %>% 
  arrange(avg_BC)

print(MeanSE_BC)

MeanSE_AD = freq_df %>%
          dplyr::group_by(temp, weight_g) %>%
          dplyr::summarise(avg_AD = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))

MeanSE_AD = MeanSE_AD %>% 
  arrange(avg_AD)

print(MeanSE_AD)

MeanSE_ABC = freq_df %>%
          dplyr::group_by(temp, time_min, depth_cm) %>%
          dplyr::summarise(avg_ABC = mean(freq),
                    se = sd(freq)/sqrt(length(freq)))

print(MeanSE_ABC)


```
### SE

```{r}


```


### plot LSD

#### temp
```{r}
p1 = ggplot(meanSE_a, aes(x = temp,
                          y = avg_a))+ 
          geom_bar(stat = "identity",
                   position = position_dodge(width=0.9),
                   width = 0.6,
                   fill = "grey")+ 
          geom_errorbar(aes(ymax = avg_a + se,
                            ymin = avg_a - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+
          labs(x = expression(paste("Temperature (",degree,"C)")),
               y = "Mean germination (%)")+
          geom_text(aes(x = temp,
                        y = avg_a + se,
                        label = as.matrix(ascend_A$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
  theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/Temperature_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)

```

#### time
```{r}
p2 = ggplot(meanSE_b, aes(x = time_min,
                          y = avg_b))+
          geom_bar(stat = "identity",
                   position = position_dodge(width=0.9),
                   width = 0.6,
                   fill = "grey")+ 
          geom_errorbar(aes(ymax = avg_b + se,
                            ymin = avg_b - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+
          labs(x = "Time (min)",
               y = "Mean germination (%)")+
          geom_text(aes(x = time_min,
                        y = avg_b + se,
                        label = as.matrix(ascend_B$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/time_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)

```

#### depht
```{r}
p3 = ggplot(meanSE_c, aes(x = depth_cm,
                          y = avg_c))+
          geom_bar(stat = "identity",
                   position = position_dodge(width=0.9),
                   width = 0.4,
                   fill = "grey")+ 
          geom_errorbar(aes(ymax = avg_c + se,
                            ymin = avg_c - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+
          labs(x = "Depth (cm)",
               y = "Mean germination (%)")+
          geom_text(aes(x = depth_cm,
                        y = avg_c + se,
                        label = as.matrix(ascend_C$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/depth_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)
```

#### weight
```{r}

p4 = ggplot(meanSE_d, aes(x = weight_g,
                          y = avg_d))+
          geom_bar(stat = "identity",
                   position = position_dodge(width=0.9),
                   width = 0.6,
                   fill = "grey")+ 
          geom_errorbar(aes(ymax = avg_d + se,
                            ymin = avg_d - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+
          geom_text(aes(x = weight_g,
                        y = avg_d + se,
                        label = as.matrix(ascend_D$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
          labs(x = "Mean sclerotial Weight (g)",
               y = "Mean germination (%)")+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/weight_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)

```

#### interaction
##### temp x time
```{r}

i1 = ggplot(MeanSE_AB, aes(x = temp,
                          y = avg_AB,
                          fill = factor(time_min)))+ 
          geom_bar(stat = "identity",
                   color = "black",
                   position = position_dodge(width=0.9))+ 
          scale_fill_grey()+ 
          geom_errorbar(aes(ymax = avg_AB + se,
                            ymin = avg_AB - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+ 
          labs(x = expression(paste("Temperature (",degree,"C)")),
               y = "Mean germination (%)",
               fill = "Time (min)")+ 
          geom_text(aes(x = temp,
                        y = avg_AB + se,
                        label = as.matrix(ascend_AB$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/temp_time_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)
```

##### temp x depth
```{r}
i2 = ggplot(MeanSE_AC, aes(x = temp,
                          y = avg_AC,
                          fill = factor(depth_cm)))+ 
          geom_bar(stat = "identity",
                   color = "black",
                   position = position_dodge(width=0.9))+ 
          scale_fill_grey(labels = c("0", "5"))+ 
          geom_errorbar(aes(ymax = avg_AC + se,
                            ymin = avg_AC - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+ 
          labs(x = expression(paste("Temperature (",degree,"C)")),
               y = "Mean germination (%)",
               fill = "Depth (cm)")+ 
          geom_text(aes(x = temp,
                        y = avg_AC + se,
                        label = as.matrix(ascend_AC$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/temp_depth_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)
```

##### time x depth
```{r}

i3 = ggplot(MeanSE_BC, aes(x = time_min,
                          y = avg_BC,
                          fill = factor(depth_cm)))+ 
          geom_bar(stat = "identity",
                   position = position_dodge(width=0.9))+ 
          scale_fill_grey(labels = c("0", "5"))+ 
          geom_errorbar(aes(ymax = avg_BC + se,
                            ymin = avg_BC - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+ 
          labs(x = "Time (min)",
               y = "Mean germination (%)",
               fill = "Depth (cm)")+ 
          geom_text(aes(x = time_min,
                        y = avg_BC + se,
                        label = as.matrix(ascend_BC$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/time_depth_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)

```


##### weight x temp

```{r}
i4 = ggplot(MeanSE_AD, aes(x = temp,
                          y = avg_AD,
                          fill = factor(weight_g)))+ 
          geom_bar(stat = "identity",
                   position = position_dodge(width=0.9))+ 
          scale_fill_grey(labels = c("0.054", "0.222", "0.670", "1.310"))+ 
          geom_errorbar(aes(ymax = avg_AD + se,
                            ymin = avg_AD - se), 
                        position = position_dodge(width=0.9), 
                        width = 0.25)+ 
          labs(x = expression(paste("Temperature (",degree,"C)")),
               y = "Mean germination (%)",
               fill = "Weight (g)")+ 
          geom_text(aes(x = temp,
                        y = avg_AD + se,
                        label = as.matrix(ascend_AD$groups)),
                    size = 6,
                    position = position_dodge(width = 0.9),
                    vjust = -(0.5))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20))+
  theme_bw()+
   theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/temp_weight_lsd.png", units = "cm", width = 35, height = 20, dpi = 300)
```


## regression tree 1

https://uc-r.github.io/regression_trees

```{r}
freq_df1 <- subset(freq_df, select=-c(size_mm))

glimpse(freq_df1)

freq_df1$weight_g <- as.factor(freq_df1$weight_g)



#data sets 

library(rsample)
set.seed(123)
split = initial_split(freq_df1, prop = 0.7)
train = training(split)
test = testing(split)

library(rpart)
library(nnet) 
library(pROC) 
library(rpart)
library(rpart.plot)

# grow the tree

m1 <- rpart(freq ~ .,freq_df1, method = "anova")

m1$cptable

png(file="C:\\Users\\2010088819\\Documents\\R Projects\\Scle_management\\plots\\tree.png",
width=600, height=350)
rpart.plot(m1, 5)  # manier 1
dev.off()


plotcp(m1)   # lyn gaan deur 4 so, gebruik 

# begin pruning m1 tree

m2 <- rpart(
  formula = freq ~ .,
  data = train,
  method = "anova",
  control = list(minsplit = 4, maxdepth = 6, xval = 10))

m1$cptable   #5


#tuning

# perform a grid search to automatically search across a range of differently tuned models to identify the optimal hyerparameter setting.

hyper_grid <- expand.grid(minsplit = seq(4, 20, 1),
                          maxdepth = seq(6, 15, 1))
                    
head(hyper_grid)

# number of combinations
nrow(hyper_grid)

# automate modelling

models <- list()

for (i in 1:nrow(hyper_grid)) {
  
# get minsplit, maxdepth values at row i
  minsplit <- hyper_grid$minsplit[i]
  maxdepth <- hyper_grid$maxdepth[i]

# train a model and store in the list
  models[[i]] <- rpart(
    formula = freq ~ .,
    data    = train,
    method  = "anova",
    control = list(minsplit = minsplit, maxdepth = maxdepth)
    )
}


# function to get optimal cp
get_cp <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  cp <- x$cptable[min, "CP"] 
}

# function to get minimum error
get_min_error <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  xerror <- x$cptable[min, "xerror"] 
}

hyper_grid %>%
  mutate(
    cp    = purrr::map_dbl(models, get_cp),
    error = purrr::map_dbl(models, get_min_error)
    ) %>%
  arrange(error) %>%
  top_n(-5, wt = error)

#apply final model and get predicted on test data

hyper_grid %>%
  mutate(
    cp    = purrr::map_dbl(models, get_cp),
    error = purrr::map_dbl(models, get_min_error)
    ) %>%
  arrange(error) %>%
  top_n(-5, wt = error)

optimal_tree <- rpart(
    formula = freq ~ .,
    data    = train,
    method  = "anova",
    control = list(minsplit = 11, maxdepth = 12, cp = 0.01)
    )

library(rattle)

png(file="C:\\Users\\2010088819\\Documents\\R Projects\\Scle_management\\plots\\optimal_tree.png",
width=600, height=350)
rpart.plot(optimal_tree, clip.right.labs = TRUE)
dev.off()


library(ipred)
library(caret)

pred <- predict(optimal_tree, newdata = test)
RMSE(pred = pred, obs = test$freq)       # 20% kan incorrect wees



# bagging
# bootstrap aggregating


# make bootstrapping reproducible
set.seed(123)

# train bagged model
bagged_m1 <- bagging(
  formula = freq ~ .,
  data    = train,
  coob    = TRUE
)

bagged_m1

# assess 10-50 bagged trees
ntree <- 10:100

# create empty vector to store OOB RMSE values
rmse <- vector(mode = "numeric", length = length(ntree))

for (i in seq_along(ntree)) {
  # reproducibility
  set.seed(123)
  
  # perform bagged model
  model <- bagging(
  formula = freq ~ .,
  data    = train,
  coob    = TRUE,
  nbagg   = ntree[i]
)
  # get OOB error
  rmse[i] <- model$err
}

plot(ntree, rmse, type = 'l', lwd = 2)+
abline(v = 25, col = "red", lty = "dashed")

# manier 2

library(caret)

# Specify 10-fold cross validation
ctrl <- trainControl(method = "cv",  number = 10) 

# CV bagged model
bagged_cv <- train(
  freq ~ .,
  data = train,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )

# assess results
bagged_cv


# plot bagged wat die most NM variables aanwys

ggplot(varImp(bagged_cv), 10)+
  geom_col(fill = "grey")+
  theme_bw()+
  labs(y = "Importance", x = "Variable")+
 theme(axis.title.x=element_text(size=18,face="bold"),
       axis.title.y=element_text(size=18,face="bold"),
       axis.text.y = element_text(size = 14, color = "black", face = "bold"),
       axis.text.x = element_text(size = 14, color = "black", face = "bold"),
       legend.text = element_text(size = 16),
       legend.title = element_text(size = 18))+
         ggsave("plots/important_variables.png", units = "cm", width = 60, height = 30, dpi = 300)

m2
```


## regression tree 2

werk nie

https://www.guru99.com/r-decision-trees.html

```{r}
# #shuffel vir wanneer train en test set maak
# 
# shuffle_index <- sample(1:nrow(freq_df))
# head(shuffle_index)
# 
# freq_df1 <- freq_df[shuffle_index, ]
# head(freq_df1)
# 
# #train set
# 
# create_train_test <- function(data, size = 0.8, train = TRUE) {
#     n_row = nrow(data)
#     total_row = size * n_row
#     train_sample < - 1:total_row
#     if (train == TRUE) {
#         return (data[train_sample, ])
#     } else {
#         return (data[-train_sample, ])
#     }
# }
# 
# # test function and dimensions
# 
# data_train <- create_train_test(freq_df1, 0.8, train = TRUE)
# data_test <- create_train_test(freq_df1, 0.8, train = FALSE)
# dim(data_train)
# 
# #grow the tree
# 
# fit <- rpart(freq ~ weight_g + temp + time_min + depth_cm,
#       data = freq_df, 
#       method = "anova")
# 
# #display results
# 
# printcp(fit)
# 
# #visualize cross validation results
# 
# plotcp(fit)
# 
# #summary of splits
# summary(fit)
# 
# # plot tree
# 
# plot(fit, uniform=TRUE, main = "Classification tree")
# 
# #create pstscript plot of tree
# 
# post(fit, file = "c:/tree.ps",
#      title = "Classification tree")

```






# dung

## import

```{r}

dung_raw = read.xlsx("C:\\Users\\2010088819\\Documents\\R Projects\\Scle_management\\manage.xlsx", "dung")


dung_raw$weight_g <- as.factor(dung_raw$weight_g)


```


##ANOVA

weight_g	num	germ_num	freq
0.030	53	2	3.773584906
0.262	24	0	0
0.328	16	2	12.5
0.400	8	0	0
0.610	3	0	0


### one way ANOVA


```{r}

# 1 - modified

weight_g <- c(0.030, 0.262, 0.328, 0.400, 0.610)
num <- c(53, 24, 16, 8, 3)
germ_num <- c(2, 0, 2, 0, 0)
freq <- c(3.77, 0, 12.5, 0, 0)


dung <- data.frame(
       Y=c(weight_g, num, freq),
       variable = factor(rep(c("weight_g", "num", "freq"),
                         times=c(length(weight_g), length(num), length(freq)))))

dung_model <- aov(Y ~ variable, data = dung)
anova(dung_model)

# 2 - raw

aov_d <- aov(germ_num ~ weight_g + num, data = dung_raw)

summary(aov_d)

write.xlsx(dung_model, "anova_dung.xlsx")

```


## graphs

```{r}


ggplot(aes(weight_g, num), data = dung_raw)+
  geom_col( fill = "grey")+
  geom_text(aes(weight_g, num, label = freq), hjust = 0.5, vjust = -0.5)+
    theme_bw()+
    labs(x = "Mean sclerotial weight (g)", y = "Number of sclerotia harvested")+
  ylim(0, 60)+
  scale_x_discrete(labels=c("0.03" = "0.030", "0.262" = "0.262",
                              "0.328" = "0.328", "0.4" = "0.400", "0.61" = "0.610"))+
  theme(axis.title.x=element_text(size=20,face="bold"),
       axis.title.y=element_text(size=20,face="bold"),
       axis.text.y = element_text(size = 16, color = "black", face = "bold"),
       axis.text.x = element_text(size = 16, color = "black", face = "bold"),
       legend.text = element_text(size = 18),
       legend.title = element_text(size = 20),
       strip.background = element_rect(fill="white", size=1, color="black"))+
         ggsave("plots/dung.png", units = "cm", width = 35, height = 20, dpi = 300)





```

, 